# qPCR analysis E. bolteae: Cecal samples from control and treatment groups
Standards:

- Standard curve: from 30,000,000 copies of inovirus to 30 copies (10-fold dilutions)
- Stock of standards comes from PCR product of the circularization regions of the inoviruses (PCR purified)

1. load the libraries
```{r}
library(tidyverse)
library(readxl)
library(ggsci)
```


2. load the data from excel
```{r}
dat <- read_excel(path = "2022May01_qPCR_cecal_Bolt_results.xlsx")
dat
```

3. create graph for standard curve and determine efficiency for E. bolt inovirus primers

## Calculate standard curve

filter `dat` so that we just have the standard data from bolt
```{r}
dat_bolt <- dat %>%
  filter(Sample == "PCR_std") %>% # keep wells with the standards
  rename(Copy_number =`Starting Quantity (SQ)`) %>%
  select(Sample, Cq, Copy_number) %>%
  group_by(Sample, Copy_number) %>%
  mutate(log_copy_number =log(Copy_number, base = 10))

dat_bolt

```


Make the standard curve for analysis
```{r}
R_squared <- summary(lm(Cq ~ log_copy_number, data = dat_bolt))$r.squared
slope <- summary(lm(Cq ~ log_copy_number, data = dat_bolt))$coefficient[2]
Efficiency <- (10^(-1/slope)-1)*100

  dat_bolt %>%
  ggplot(aes(x = Copy_number, y = Cq)) +
    geom_point(color = "red") +
    geom_smooth(method = "lm", se = F, color = "black") + 
    annotate("text",x = 1e+06, y = 25, label =  paste0("R^2= ", round(R_squared, digits = 4))) +
    scale_x_log10() +
    annotate("text",x = 1e+06, y = 24, label =  paste0("Efficiency= ",round(Efficiency, digits = 2), "%"), color = "red") +
  
    theme_bw() +
    labs(x = "Inovirus Copies/Rxn", y = "Cq", title = "C. boltea Inovirus Standard Curve")

```

## Calculate the copy number of samples
Add the Weight of cecal sample and dilution factor (DF) to the rows

Load Dilution factor values and weight of the cecal sample
```{r}
map_file <- read_excel(path = "2022May01_qPCR_cecal_Bolt_results.xlsx", sheet = "Sheet2") %>%
  mutate(Sample = substring(Sample, 1,4), Cage_num = factor(Cage_num)) %>%
  mutate(Cage_num = paste0("Cage #", Cage_num)) %>%
  mutate(Treatment_PEG = paste0(Treatment_PEG, "% PEG"))

map_file
```

Make data frame with predicted values for inovirus.
```{r}
# Linear regression model -> we will use this to interpolate the samples' Cq values
Bolt_lm <- lm(Cq ~ log_copy_number, data = dat_bolt)
slope <- coefficients(Bolt_lm)[2]
intercept <- coefficients(Bolt_lm)[1]


dat_samples<- dat %>%
  filter(!Sample %in% c("PCR_std", "gDNA_std"), Target == "Inov") %>%
  mutate(Target = factor(Target), Sample = factor(Sample)) %>%
  full_join(map_file) %>%
  
  select(Sample, Strain, Cq, Cage_num, Weight_sample, DF, Treatment_PEG) %>%
  mutate(Copy_number = round(10^((Cq -intercept)/slope),0)) %>%
  mutate(Copy_number_g = Copy_number*6*DF/Weight_sample) %>%
  
  # # Calculate the mean copy number for each sample
  group_by(Sample, Strain, Cage_num, Treatment_PEG) %>%
  summarise(mean_Copy_number = mean(Copy_number),
            mean_Copy_number_g = mean(Copy_number_g))

dat_samples
```


Look at the Cq values for samples
```{r}
dat %>%
  filter(!Sample %in% c("PCR_std", "gDNA_std"), Target == "Inov") %>%
  mutate(Target = factor(Target), Sample = factor(Sample)) %>%
  ggplot(aes(x = Sample, y = Cq)) +
    geom_boxplot() +
    scale_y_continuous(limits = c(0,30), breaks = seq(0,28,2)) +
    theme_classic() +
  scale_color_jama()
```


plot experimental samples

```{r}
dat_samples %>%
  ggplot(aes(x = Treatment_PEG, y = mean_Copy_number_g)) +
  geom_boxplot(aes(fill = Treatment_PEG, color = Treatment_PEG), alpha = 0.3, width = 0.5) +
  geom_point(width = 0.1, size =1.5) +
  scale_y_log10(limits = c(1e+04, 1e+10), n.breaks = 6) +
  scale_color_npg() +
  theme_classic()

```


# genome copy number qPCR for E. boltea

## Calculate standard curve

filter `dat` so that we just have the standards data from bolt (gDNA)
```{r}
dat_genome_std <- dat %>%
  filter(Sample == "gDNA_std") %>%
  rename(Copy_number =`Starting Quantity (SQ)`) %>%
  filter(Copy_number > 100) %>% # Take out this as it did not amplify
  select(Sample, Cq, Copy_number) %>%
  group_by(Sample, Copy_number) %>%
  mutate(log_copy_number =log(Copy_number, base = 10))

dat_genome_std
```


Make the standard curve for analysis
```{r}
R_squared <- summary(lm(Cq ~ log_copy_number, data = dat_genome_std))$r.squared
slope <- summary(lm(Cq ~ log_copy_number, data = dat_genome_std))$coefficient[2]
Efficiency <- (10^(-1/slope)-1)*100

  dat_genome_std %>%
  ggplot(aes(x = Copy_number, y = Cq)) +
    geom_point(color = "red") +
    geom_smooth(method = "lm", se = F, color = "black") + 
    annotate("text",x = 1e+06, y = 25, label =  paste0("R^2= ", round(R_squared, digits = 4))) +
    scale_x_log10() +
    annotate("text",x = 1e+06, y = 24, label =  paste0("Efficiency= ",round(Efficiency, digits = 2), "%"), color = "red") +
  
    theme_bw() +
    labs(x = "Inovirus Copies/Rxn", y = "Cq", title = "C. boltea gDNA Standard Curve")

```

Make data frame with predicted values
```{r}
# Linear regression model
Bolt_genome_lm <- lm(Cq ~ log_copy_number, data = dat_genome_std)
slope <- coefficients(Bolt_genome_lm)[2]
intercept <- coefficients(Bolt_genome_lm)[1]


dat_samples_gDNA<- dat %>%
  
  # Filter data you want and then calculate copy number for each sample
  filter(!Sample %in% c("PCR_std", "gDNA_std"), Target == "gDNA") %>%
  mutate(Target = factor(Target), Sample = factor(Sample)) %>%
  full_join(map_file) %>%
  
  select(Sample, Strain, Cq, Cage_num, Weight_sample, DF, Treatment_PEG) %>%
  mutate(Copy_number_gDNA = round(10^((Cq -intercept)/slope),0)) %>%
  mutate(Copy_number_gDNA_g = Copy_number_gDNA*6*DF/Weight_sample) %>%
  
  # # Calculate the mean copy number for each sample
  group_by(Sample, Strain, Cage_num, Treatment_PEG) %>%
  summarise(mean_Copy_number_gDNA = mean(Copy_number_gDNA),
            mean_Copy_number_gDNA_g = mean(Copy_number_gDNA_g))

dat_samples_gDNA
```

check the Cq values
```{r}
dat %>%
  filter(!Sample %in% c("PCR_std", "gDNA_std"), Target == "gDNA") %>%
  mutate(Target = factor(Target), Sample = factor(Sample)) %>%
  ggplot(aes(x = Sample, y = Cq)) +
    geom_boxplot() +
    geom_point()+
    scale_y_continuous(limits = c(10,25), breaks = seq(10,25,2)) +
    theme_classic() +
  scale_color_jama()
```

```{r}
dat_samples_gDNA %>%
  ggplot(aes(x = Treatment_PEG, y = mean_Copy_number_gDNA_g)) +
  geom_boxplot(aes(fill = Treatment_PEG, color = Treatment_PEG), alpha = 0.3, width = 0.5) +
  geom_point(width = 0.1, size =1.5) +
  scale_y_log10(limits = c(1e+07, 1e+12), n.breaks = 6) +
  scale_color_npg() +
  theme_classic()
```

# Joing the two data sets and plot
```{r}
norm_dat_Bolt <- dat_samples %>%
  select(Sample, Strain, Cage_num,Treatment_PEG, mean_Copy_number, mean_Copy_number_g) %>%
  full_join(dat_samples_gDNA %>% select(Sample, Strain, Cage_num, mean_Copy_number_gDNA, mean_Copy_number_gDNA_g)) %>%
  mutate(Norm_copy_number = round((mean_Copy_number/mean_Copy_number_gDNA)*1000000, 0))

norm_dat_Bolt
```

plot experimental samples
```{r}
norm_dat_Bolt %>%
  ggplot(aes(x= Treatment_PEG, y= Norm_copy_number)) +
  stat_summary(geom = "errorbar", fun.data = "mean_sdl", 
            fun.args = list(mult = 1), width = 0.1) +
  stat_summary(geom = "bar", aes(fill = Treatment_PEG), fun.y = mean, color = "black", alpha = 0.5) +
  geom_jitter(width = 0.01, size =2, alpha = 0.8) +
  scale_y_continuous(n.breaks = 12) + 
  scale_fill_aaas() +
  theme_classic()
 
```

