
# correlation plots

load the libraries
```{r}
library(tidyverse)
library(readxl)
library(ggsci)
library(ggpubr)
```


load the data:
```{r}
dat <- read_excel("osmo_compiled_experiment.xlsx") %>%
  mutate(Bacteria = factor(Bacteria))
dat

dat$Bacteria <- factor(dat$Bacteria, levels = c("Enterocloster bolteae",
                                                "Enterocloster clostridioformis 538",
                                                "Enterocloster clostridioformis 455",
                                                "Enterocloster citroniae"))
```


# Plot for growth rate vs inovirus copy number
```{r}
breaks_log10 <- function(x) {
  low <- floor(log10(min(x)))
  high <- ceiling(log10(max(x)))

  10^(seq.int(low, high, by = 2))
}

p <- dat %>%
  ggplot(aes(x = Growth_Rate, y = `Inovirus pI copies/mL supernatant`, color = Bacteria)) +
    geom_point() +
    geom_smooth(method = "lm")+

      scale_y_log10(breaks = breaks_log10, limits = c(1e+3, 1e+8),n.breaks = 6, expand = c(0, 0),
                labels = scales::trans_format(log10, scales::math_format(10^.x))) +
    annotation_logticks(sides = "l") +
    scale_x_continuous(limits = c(0, 0.30), breaks = seq(0,0.30,0.05), expand = expansion(mult = 0)) +
    scale_color_aaas() +
    theme_light() +
    labs(x = "Growth Rate (OD/h)", y = "Inovirus pI copies/mL", color = "Strain") +
    theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "white"),
        axis.ticks = element_line(color = "black")) 
p


```
# Look at the correlation coefficient value for each correlation

## 1) C. bolteae assumptions and correlation test
```{r}
dat_bolt <- dat %>% filter(Bacteria == "Enterocloster bolteae")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_bolt$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_bolt$`Inovirus pI copies/mL supernatant`)

res <- cor.test(dat_bolt$Growth_Rate, dat_bolt$`Inovirus pI copies/mL supernatant`, method = "spearman")
res
```
## 2) C. clostridioformis 538 assumptions and correlation test
```{r}
dat_EC538 <- dat %>% filter(Bacteria == "Enterocloster clostridioformis 538")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_EC538$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_EC538$`Inovirus pI copies/mL supernatant`)

# We need to use spearman, since once of the variables is not normally distributed!!
res <- cor.test(dat_EC538$Growth_Rate, dat_EC538$`Inovirus pI copies/mL supernatant`, method = "spearman")
res
```

## 3) C. clostridioformis 455 assumptions and correlation test
```{r}
dat_EC455 <- dat %>% filter(Bacteria == "Enterocloster clostridioformis 455")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_EC455$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_EC455$`Inovirus pI copies/mL supernatant`)

# We need to use spearman, since once of the variables is not normally distributed!!
res <- cor.test(dat_EC455$Growth_Rate, dat_EC455$`Inovirus pI copies/mL supernatant`, method = "spearman")
res
```
## 4) C. citroniae assumptions and correlation test
```{r}
dat_citroniae <- dat %>% filter(Bacteria == "Enterocloster citroniae")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_citroniae$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_citroniae$`Inovirus pI copies/mL supernatant`)

# We need to use spearman, since once of the variables is not normally distributed!!
res <- cor.test(dat_citroniae$Growth_Rate, dat_citroniae$`Inovirus pI copies/mL supernatant`, method = "spearman")
res
```

# Biofilm formations correlation with Inovirus pI copy numbers

```{r}


p <- dat %>%
  ggplot(aes(x = `Biofilm Formation OD 595`, y = `Inovirus pI copies/mL supernatant`, color = Bacteria)) +
    geom_point() +
    geom_smooth(method = "lm") +

      scale_y_log10(breaks = breaks_log10, limits = c(1e+3, 1e+8),expand = c(0, 0), n.breaks = 8,
                labels = scales::trans_format(log10, scales::math_format(10^.x))) +
    annotation_logticks(sides = "l") +
    scale_x_continuous(limits = c(0, 3.2), breaks = seq(0,3.5,0.5), expand = expansion(mult = 0)) +
    
  #facet_grid(.~Bacteria, scales = "free")+
  scale_color_aaas() +
    theme_light() +
    labs(x = "Biofilm formation (OD595)", y = "Inovirus pI copies/mL", color = "Strain") +
    theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "white"),
        axis.ticks = element_line(color = "black")) 
p

```

# Look at the R value for each correlation

## 1) C. bolteae assumptions and correlation test
```{r}
dat_bolt <- dat %>% filter(Bacteria == "Enterocloster bolteae")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_bolt$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_bolt$`Biofilm Formation OD 595`)

res <- cor.test(dat_bolt$Growth_Rate, dat_bolt$`Biofilm Formation OD 595`, method = "spearman")
res
```
## 2) C. clostridioformis 538 assumptions and correlation test
```{r}
dat_EC538 <- dat %>% filter(Bacteria == "Enterocloster clostridioformis 538")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_EC538$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_EC538$`Biofilm Formation OD 595`)

# We need to use spearman, since once of the variables is not normally distributed!!
res <- cor.test(dat_EC538$Growth_Rate, dat_EC538$`Biofilm Formation OD 595`, method = "spearman")
res
```

## 3) C. clostridioformis 455 assumptions and correlation test
```{r}
dat_EC455 <- dat %>% filter(Bacteria == "Enterocloster clostridioformis 455")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_EC455$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_EC455$`Biofilm Formation OD 595`)

# We need to use spearman, since once of the variables is not normally distributed!!
res <- cor.test(dat_EC455$Growth_Rate, dat_EC455$`Biofilm Formation OD 595`, method = "spearman")
res
```
## 4) C. citroniae assumptions and correlation test
```{r}
dat_citroniae <- dat %>% filter(Bacteria == "Enterocloster citroniae")

# Shapiro-Wilk normality test for Growth Rate
shapiro.test(dat_citroniae$Growth_Rate) 
# Shapiro-Wilk normality test Inovirus copy number
shapiro.test(dat_citroniae$`Biofilm Formation OD 595`)

# We need to use spearman, since once of the variables is not normally distributed!!
res <- cor.test(dat_citroniae$Growth_Rate, dat_citroniae$`Biofilm Formation OD 595`, method = "spearman")
res
```






