#1. load library
```{r}
library(tidyverse)
library(ggsci)
```


#2. read data
```{r}
dat <- read_csv("genomes_metadata_full.csv") %>%
  mutate(genome_name = taxonomy)  %>%
  select(taxonomy, genome_name, dplyr::everything())
dat
```
#3. replace spaces and special characters with '_' 
```{r}

dat_new <- dat %>% 
  mutate(genome_name = gsub("/", replacement = "_", genome_name, fixed = T)) %>%
  mutate(genome_name = gsub("-", replacement = "_", genome_name, fixed = T)) %>%
  mutate(genome_name = gsub(".", replacement = "", genome_name, fixed = T)) %>%
  mutate(genome_name = gsub("  ", replacement = "_", genome_name)) %>%
  mutate(genome_name = gsub(" ", replacement = "_", genome_name))

dat_new
```
#4. save as a .tsv file to be used in sockeye
```{r}
#write_tsv(x = dat_new, file = "genome_list.tsv")
```

#5. Now add a `yes` to the strains that have a predicted inovirus and a `no` to those that dont
- From our analysis, we found that only the 5 clostridia strains had predicted inoviruses, the other ones did not

```{r}
# names of strains with inovirus predictions
predictions <- c("Clostridium_aldenense_AM40_2AC",
                 "Clostridium_bolteae_ATCC_BAA_613",
                 "Clostridium_bolteae_WAL_14578_BEI_HM_318",
                 "Clostridium_citroniae_WAL_17108_BEI_HM_315",
                 "Clostridium_clostridioforme_2_1_49FAA_BEI_HM_306",
                 "Clostridium_clostridioforme_WAL_7855_BEI_HM_317",
                 "Clostridium_hathewayi_WAL_18680_BEI_HM_308")

# add a column names predicted_inov
dat_inov <- dat_new %>%
  mutate(predicted_inov = if_else(genome_name %in% predictions, true = "Yes", false = "No"))
dat_inov
```
# 6. Add the names that will be plotted in the x-axis of the plot
```{r}
names <-dat_inov %>%
  group_by(genus) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  dplyr::select(genus) %>%
  pull
names

dat_inov$genus <- factor(dat_inov$genus, levels = names)
```

# 7. Plot the inovirus data
```{r}
p <- dat_inov %>%
  #mutate(predicted_inov = factor(predicted_inov), predicted_inov = fct_inorder(predicted_inov)) %>%
  ggplot(aes(x = genus, fill = predicted_inov)) +
    geom_bar(color = "black", width = 0.8) +
    scale_fill_manual(values = c("grey20", "red")) +
    scale_y_continuous(limits = c(0,40), breaks = seq(0,40,2), expand = c(0,0)) +
  theme_light() +
  labs(fill = "Inovirus Predicted?") +
    theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "white"),
        axis.ticks = element_line(color = "black"),
        #strip.text.x = element_text(size=12, face="italic", color = "black"),
        strip.text.x = element_blank(),
        axis.title = element_text(size = 12),
        legend.position = c(0.90, 0.8),
        axis.text=element_text(size=16),
        axis.text.x=element_text(angle = 50, hjust = 1, vjust = 1)
    )
p
```

Pie chart of the different phyla in the data frame
```{r}
genus <- dat_inov %>%
  group_by(phylum) %>%
  count() %>%
  ungroup() %>%
  select(1) %>%
  pull

values <- dat_inov %>%
  group_by(phylum) %>%
  count() %>%
  ungroup() %>%
  select(2) %>%
  pull

pie(values, genus)
```













