# CFU per mL analysis of fecal and cecal samples for Inovirus mouse experiment

#1 .load the libraries for the analysis  
```{r}
library(tidyverse)
library(readxl)
library(ggsci)
```


#2. load the data to plot
```{r}
dat <- read_excel(path = "CFU_per_mL.xlsx", sheet = "Sheet1") %>%
  mutate(Media = factor(Media), Media = fct_rev(Media)) %>%
  mutate(Cage_num = factor(Cage_num),
         Treatment_PEG = factor(Treatment_PEG),
         Cage_num = fct_relabel(Cage_num, .fun = function(.) paste0("Cage #", .)),
         Treatment_PEG = fct_relabel(Treatment_PEG, .fun = function(.) paste0(.,"% PEG")))

dat
```

#3 Plot CFU_mL -> fecal
```{r}
dat %>%
  filter(Sample_type == "Fecal") %>%
  ggplot(aes(x = Strain, y = CFU_mL, fill = Treatment_PEG)) +
    geom_boxplot() +
    geom_point(position=position_jitterdodge()) +
    scale_y_log10(limits = c(1e+0, 1e+13), n.breaks = 13 ,expand = c(0, 0)) +
    theme_classic() +
    scale_fill_npg() +
    labs(x = "Strain", y = "CFU/mL", fill = "PEG in water (w/v)") +
    facet_grid(.~Media) +
  
    theme(strip.text.x = element_text(size=10, colour = "Black"),
        strip.background = element_rect(colour="black", fill="grey90"),
        axis.text.x = element_text(angle = 45,hjust=1, face = "italic")
        #panel.grid = element_blank()
        )
```




#3 Plot CFU_mL -> cecal
```{r}
dat %>%
  filter(Sample_type == "Cecal") %>%
  ggplot(aes(x = Strain, y = CFU_mL, fill = Treatment_PEG)) +
    geom_boxplot() +
    geom_point(position=position_jitterdodge()) +
    scale_y_log10(limits = c(1e+0, 1e+13), n.breaks = 13 ,expand = c(0, 0)) +
    theme_classic() +
    scale_fill_npg() +
    labs(x = "Strain", y = "CFU/mL", fill = "PEG in water (w/v)") +
    facet_grid(.~ Media) +
  
    theme(strip.text.x = element_text(size=10, colour = "Black"),
        strip.background = element_rect(colour="black", fill="grey90"),
        axis.text.x = element_text(angle = 45,hjust=1, face = "italic")
        #panel.grid = element_blank()
        )
```


#3 Plot CFU_mL -> cecal
```{r}
breaks_log10 <- function(x) {
  low <- floor(log10(min(x)))
  high <- ceiling(log10(max(x)))

  10^(seq.int(low, high, by = 2))
}

p <- dat %>%
  filter(Sample_type == "Cecal", Media == "Columbia blood") %>%
  ggplot(aes(x = Strain, y = CFU_mL, fill = Treatment_PEG)) +
    geom_boxplot(alpha=0.5) +
    geom_point(position=position_jitterdodge()) +
      facet_grid(.~ Media) +
    #scale_y_log10(limits = c(1e+0, 1e+13), n.breaks = 13 ,expand = c(0, 0)) +

    scale_y_log10(breaks = breaks_log10, limits = c(1e+0, 1e+12),expand = c(0, 0), n.breaks = 6,
                labels = scales::trans_format(log10, scales::math_format(10^.x))) +
    annotation_logticks(sides = "l") +
  scale_x_discrete()+
    theme_light() +
    scale_fill_aaas() +
    labs(x = "Strain", y = "Bacterial counts (CFU/mL)", fill = "PEG in water (w/v)") +

  
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "grey96"),
        axis.ticks = element_line(color = "black"),
        #strip.text.x = element_text(size=12, face="italic", color = "black"),
        strip.text.x = element_blank(),
        axis.title = element_text(size = 12),
        axis.text=element_text(size=16, face = "italic"),
        legend.position = c(0.75, 0.2)
        ) 
p
```

# 4 test for significance using t-test

Test for E. bolteae: Calculate significance using two sample t-test
```{r}
# filter dat
dat_bolt <- dat %>%
  filter(Strain == "E. bolteae")
dat_bolt

# do barlett test to check for equal variance
bartlett.test(CFU_mL~Treatment_PEG, data = dat_bolt) # use Var.equal = FALSE

# Perform two sample t test
t.test(CFU_mL~Treatment_PEG, data = dat_bolt,
       var.equal = FALSE,
       conf.level = 0.95)

```

Test for E. clostridioformis 538: Calculate significance using two sample t-test
```{r}
# filter dat
dat_CF538 <- dat %>%
  filter(Strain == "E. clostridioformis 538")
dat_CF538

# do barlett test to check for equal variance
bartlett.test(CFU_mL~Treatment_PEG, data = dat_CF538) # use Var.equal = FALSE

# Perform two sample t test
t.test(CFU_mL~Treatment_PEG, data = dat_CF538,
       var.equal = FALSE,
       conf.level = 0.95)

```




