
# Visualizing the osmolality data

## 1. load the libraries
```{r}
library(tidyverse)
library(readxl)
library(ggsci)
```

## 2.load the data:
```{r}
dat <- read_excel("osmo_compiled_experiment.xlsx") %>%
  mutate(Bacteria = factor(Bacteria))
dat

dat$Bacteria <-factor(dat$Bacteria, levels = c("Enterocloster bolteae",
                                               "Enterocloster clostridioformis 538",
                                               "Enterocloster clostridioformis 455", 
                                               "Enterocloster citroniae"))
```

## Make a Plot to compare MaxOD600 compared osmolality
```{r}

p <- dat %>%
  ggplot(aes(x = Osmolality, y = Max_OD, colour = Bacteria)) +
    stat_summary(geom = "errorbar", fun.data = "mean_sdl", 
            fun.args = list(mult = 1), width = 5, size = 0.8) +
    stat_summary(geom = "line", fun = mean, size = 1, alpha = 0.5) +
    stat_summary(geom = "point", fun = mean, size = 2) +
    #geom_point(alpha = 0.9, size = 2) +

  scale_y_continuous(limits = c(0, 1.6), n.breaks = 8, expand = expansion(mult = 0)) +
  scale_color_aaas() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "Max OD600", color = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "white"),
        axis.ticks = element_line(color = "black"),
        legend.position = c(0.25, 0.25)) 
p
```

## Make a Plot to compare growth rate compared osmolality
```{r}
p2 <- dat %>%
  ggplot(aes(x = Osmolality, y = Growth_Rate, colour = Bacteria)) +
    stat_summary(geom = "errorbar", fun.data = "mean_sdl", 
            fun.args = list(mult = 1), width = 5, size = 0.8) +
    stat_summary(geom = "line", fun = mean, size = 1, alpha = 0.5) +
    #stat_summary(geom = "point", fun = mean, size = 2) +
    geom_point(alpha = 0.9, size = 2) +

  scale_y_continuous(limits = c(0, 0.30), breaks = seq(0,0.30,0.05), expand = expansion(mult = 0)) +
  scale_color_aaas() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "Growth Rate (OD/h)", color = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "white"),
        axis.ticks = element_line(color = "black")) 
p2
```

## Plot for MaxOD600 - boxplot
```{r}
p <- dat %>%
  ggplot(aes(x = Osmolality, y = Max_OD, group = interaction(Osmolality, Bacteria), fill = Bacteria)) +
    geom_boxplot(alpha = 0.8) +
    #stat_summary(geom = "point", fun = mean, size = 2) +
    #geom_point( position=position_jitterdodge()) +

  scale_y_continuous(limits = c(0, 1.6), n.breaks = 8, expand = expansion(mult = 0)) +
  scale_fill_aaas() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "Max OD600", color = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "white"),
        axis.ticks = element_line(color = "black"),
        legend.position = c(0.25, 0.25)) 
p
```
## Plot for growth rate - boxplots
```{r}

p2 <- dat %>%
  ggplot(aes(x = Osmolality, y = Growth_Rate, group = interaction(Osmolality, Bacteria), fill = Bacteria)) +
    geom_boxplot(alpha = 0.8) +


  scale_y_continuous(limits = c(0, 0.18), n.breaks = 9, expand = expansion(mult = 0)) +
  scale_fill_jama() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "Growth Rate (OD/h)", color = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "white"),
        axis.ticks = element_line(color = "black"),
        legend.position = c(0.25, 0.25)) 
p2
```


## plot - growth rate against osmolality (Facet by strain)
```{r}
p <- dat %>%
  ggplot(aes(x = Osmolality, y = Growth_Rate, group = interaction(Osmolality, Bacteria), fill = Bacteria)) +
    geom_boxplot(alpha= 0.5) +
    facet_grid(.~ Bacteria) +

  scale_y_continuous(limits = c(0, 0.30), breaks = seq(0,0.30, 0.05)) +
  scale_x_continuous(expand = expansion(add = 20)) +
  scale_fill_aaas() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "Growth Rate (OD/h)", fill = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "grey96"),
        axis.ticks = element_line(color = "black"),
        #strip.text.x = element_text(size=12, face="italic", color = "black"),
        strip.text.x = element_blank(),
        strip.background = element_blank()
        ) 
p
```

## plot - MaxOD600 against osmolality (Facet by strain)
```{r}

p <- dat %>%
  ggplot(aes(x = Osmolality, y = Max_OD, group = interaction(Osmolality, Bacteria), fill = Bacteria)) +
    geom_boxplot(alpha= 0.5) +
    facet_grid(.~ Bacteria) +

  #scale_y_continuous(limits = c(0, 0.30), breaks = seq(0,0.30, 0.05)) +
  scale_y_continuous(limits = c(0, 1.6), n.breaks = 8, expand = c(0,0)) +
  scale_x_continuous(expand = expansion(add = 20)) +
  scale_fill_aaas() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "Max OD600", color = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "grey96"),
        axis.ticks = element_line(color = "black"),
        #strip.text.x = element_text(size=12, face="italic", color = "black"),
        axis.text=element_text(size=16),
        strip.text.x = element_blank()
        ) 
p
```


## plot - inovirus copies against osmolality (Facet by strain)
```{r}
breaks_log10 <- function(x) {
  low <- floor(log10(min(x)))
  high <- ceiling(log10(max(x)))

  10^(seq.int(low, high))
}


p <- dat %>%
  ggplot(aes(x = Osmolality, y = `Inovirus pI copies/mL supernatant`, group = interaction(Osmolality, Bacteria), fill = Bacteria)) +
    geom_boxplot(alpha= 0.5) +
    facet_grid(.~ Bacteria) +
  annotation_logticks() +

    scale_y_log10(breaks = breaks_log10,
                labels = scales::trans_format(log10, scales::math_format(10^.x)), expand = c(0,0)) +
  scale_x_continuous(expand = expansion(add = 20)) +
  scale_fill_aaas() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "Inovirus pI copies/mL supernatant", color = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "grey96"),
        axis.ticks = element_line(color = "black"),
        #strip.text.x = element_text(size=12, face="italic", color = "black"),
        strip.text.x = element_blank(),
        axis.text=element_text(size=16),
        strip.background = element_blank()
        ) 
p
```


## plot - biofilm formation against osmolality (Facet by strain)
```{r}
p <- dat %>%
  ggplot(aes(x = Osmolality, y = `Biofilm Formation OD 595`, group = interaction(Osmolality, Bacteria), fill = Bacteria)) +
    geom_boxplot(alpha= 0.5) +
    facet_grid(.~ Bacteria) +

  scale_y_continuous(limits = c(0, 3), breaks = seq(0,3, 0.5)) +
  scale_x_continuous(expand = expansion(add = 20)) +
  scale_fill_aaas() +
  theme_light() +
  labs(x = "Osmolality (mOsm/kg)", y = "OD595", fill = "Strain") +
  theme(panel.border = element_rect(color = "black"),
        panel.grid = element_line(colour = "grey96"),
        axis.ticks = element_line(color = "black"),
        #strip.text.x = element_text(size=12, face="italic", color = "black"),
        strip.text.x = element_blank(),
        strip.background = element_blank()
        ) 
p
```















